|

### 2022-05-17 Memo/memo.Rmd
small changes to memo and appendix

We agree with Reviewer 1 that we needed to provide more information about our
online survey. Therefore, we have detailed our survey procedures in Section A.14
of the Supplementary Materials (pages 22-23), and we also briefly discuss our
survey methodology on pages 17 and 22 (footnote 13) of the main article.
Furthermore, we have rewritten paragraph 2 of Section 6.2 to clarify the
importance of the survey results in our paper. It now reads as follows:
``\textit{We assume that city councilors want to extract rents and to get
reelected. The rents come from continuing in the repeated game, and the
partisanship discounts that benefit. Figure 2 shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition. This
provides further evidence that mayors do offer city councilors the benefits they
demand. In this regard, granting favors works as rent transfers for councilors.
Job appointments are also convenient for the mayors because patronage in Brazil
has positive results in terms of welfare \citep{toral2019benefits}.}''

>>>>

We agree with Reviewer 1 that we needed to provide more information about our
online survey. Therefore, we have detailed our survey procedures in Section A.14
of the Supplementary Materials (pages 23-24), and we also briefly discuss our
survey methodology on pages 17 and 22 (footnote 13) of the main article.
Furthermore, we have rewritten paragraph 2 of Section 6.2 to clarify the
importance of the survey results in our paper. It now reads as follows:
``\textit{We assume that city councilors want to extract rents and to get
reelected. The rents come from continuing in the repeated game, and the
partisanship discounts that benefit. Figure 2 shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition. This
provides further evidence that mayors do offer city councilors the benefits they
demand. In this regard, granting favors works as rent transfers for councilors.
Job appointments are also convenient for the mayors because patronage in Brazil
has positive results in terms of welfare \citep{toral2019benefits}.}''

### 2022-05-17 appendix/appendix.Rmd
update memo and appendix
small changes to memo and appendix

In this section, we show the proofs for the model included in the article.

@@

\begin{proof}
To prove this result, we need to show that the mayor's expected utility satisfies the increasing differences in $g$ and $N$. This means that, when increasing the size of the council, the optimal solution $g^*$ should also increase. However, as $N$ increases discretely, we cannot take the cross-partial derivative on $N$ and $g$ or use the implicit function theorem. Instead, we use monotone comparative statics to derive these results \citep{milgrom1994monotone}.

@@

Therefore, costs always increase when legislature size increases.

@@

\begin{itemize}
\item \textbf{Number of Seats 2000}: Number of city councilors in 2000 (source: TSE).
\item \textbf{Population 2000}: City population according to the 2000 Brazilian Census (source: IBGE).
\item \textbf{Per Capita GDP 2000 (Millions)}: Municipal per capita GDP measured by the Brazilian Census (source: IPEA).
\item \textbf{Proportion of Poverty 2000}: Percentage of citizens earning less than R\$ 70.00 per month (source: Social Development Ministry [MDS]).
\item \textbf{Number of Seats 2004}: Number of city councilors in 2004 according to the Electoral Court (source: TSE).
\item \textbf{Infant Mortality 2005--2008}: Number of children who died before reaching one year of age, divided by the number of children born alive, multiplied by 1,000 (source: DataSUS)
\item \textbf{Postneonatal Mortality 2005--2008}: Number of children alive for at least 28 days who died before reaching one year of age, divided by the total number of children who lived for 28 days, multiplied by 1,000 (source: DataSUS)
\item \textbf{Enrollment Elementary School 2005--2008}: Number of children enrolled in elementary schools, averaged by classroom size (K--4) (source: INEP).
\item \textbf{Quality of Elementary School 2005--2008}: Average municipal IDEB scores. IDEB scores are estimated as follows. The Ministry of Education takes the average of the students' grades in math and Portuguese, then multiply the those numbers by the student yearly approval rate. INEP reweights the estimators to avoid school and classroom specific distortions (source: INEP, years: 2005, 2007).
\item \textbf{Size of Mayoral Pre-Electoral Coalition 2004}: Number of elected councilors in the mayoral pre-electoral coalition (source: TSE)
\item \textbf{Number of Appointed Bureaucrats 2005--2008}: Number of bureaucrats of the direct administration who were appointed to jobs in the municipality (source: IBGE, years: 2005, 2006, and 2008)
\item \textbf{Number of Females Elected 2004}: Number of females elected to the city council (source: TSE)
\item \textbf{Number of Non-Whites Elected 2004}: Number of non-whites elected to the city council. We collect these data for municipalities less than 10,000 inhabitants away from the cutoffs (source: own compilation based on the TSE candidate pictures).
\item \textbf{Competition per Seat 2004}: Number of people running for city councilor divided by the city council size (source: TSE)
\item \textbf{Proportion Approved Legislation 2005}: Number of approved legislation in 2005 divided by the number of proposed legislation (source: InterLegis)
\end{itemize}

@@

The inclusion of controls in our model has another purpose, which is to improve the efficiency of our estimates \citep{calonico2019regression}. We add five controls to the main specifications. We include the 2003 population because it conditions the multiple threshold assignments. We add GDP per capita to control for Brazil's economic expansion during the 2000s. We also add the number of seats in 2000 to ensure that our results come from the current legislature size and not from past seat allocations. Year was included in the models to control for seasonal effects. Finally, we add a dummy for Northeast states. The Northeast region received considerable attention from the Federal Government in the 2000s. According to \citet{de2015politica}, in 2003 the Lula administration recreated the Superintendency for the Development of the Northeast (SUDENE), investing BRL 84 Billion from 2003 to 2010. This amounts to around 70% of all targeted regional investments. One example of these investments is that the Federal Government created about 100 new universities in the Northeast region. Moreover, infant mortality in the region decreased from 22.4 deaths per 1000 infants born in 2005 to 19.6 deaths per 1000 infants born in 2008 due to larger federal investments. In turn, education quality improved from 2.69 in 2005 to a 3.14 score in 2007 (0 to 10 quality of education score). For this reason, adding a dummy for Northeast states is essential for ensuring the consistency and the efficiency of our estimates.

@@

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{0.9\linewidth}
      \centerline{\includegraphics[height=3.45in,angle=0]{../figures/simdistrRDD.pdf}}
      \caption{Simulation: RD with Multiple Thresholds -- Data Distribution}
      \bigskip
      \label{simRDdd}
    \end{minipage}
  \end{center}
\end{figure}

@@

@@

For instance, changes in controls substantially affect the quality of elementary education, where some sets of control variables leading to a positive effect. However, most combinations of controls were insignificant.

@@

Here we evaluate the robustness of our results when we change the number of cutoffs. We run a series of regressions limiting the population sizes to half the cutoffs, starting from the second cutoff to the last. This strategy is similar to adding one cutoff at a time. The results are stable and do not change our main conclusions. Moreover, there are no clear increasing or decreasing patterns, which would indicate the existence of differential returns. Figures \ref{senscut1} and \ref{senscut2} display the results for the main cutoffs.

@@

Two alternative explanations may confound our results. First, the mayor's selection could be affected by the size of the city council. Thus, public welfare can improve due to the quality of the mayor and not because of the partisan mechanism we describe. Second, the partisan mechanism could be capturing a dynamic similar to @Weingastetal1981 "law of 1/n", whereby welfare improvements are caused by higher available revenues. This section shows that none of these concerns are valid in our case.

@@

We hand-coded 1\% of the dataset (3,466 cases) and applied a Supporting Vector Machine (SVM) algorithm to the remaining 99\% cases to classify the legislation. First, we trained and tested the accuracy of the SVM classifier on 80\% of the dataset. Then, we ran the training algorithm on the complete hand-coded data and predicted the remaining cases.\footnote{We tested SVM, Naïve Bayes, Random Forests, and Neural Networks for this task. We have chosen SVM as it gives the highest prediction rate of all of the algorithms we have tried. We use a simple bag-of-words classifier provided by the `RTextTools` package for `R` \citep{collingwood2013rtexttools}.}

@@

To weight the survey, we use the Legislative Census, conducted by *Interlegis* on the behalf of the Brazilian Senate. The company was hired to improve the quality of Brazilian government branches. We include the following categories in the weighting procedure: region of Brazil, legislature size (9 to 15 or more), gender, and age less than 39. We use the proportions in each bin to weight the graphs and statistics generated after the raking process.\footnote{The `survey` package for `R` describes the raking process in more detail \citep{lumley2004analysis}. In summary, it iterates the post-stratification procedure until the sample marginals match the population marginals for all variables.} Table \ref{tabweight} shows the sample, population, and weighted proportions.

@@

\begin{enumerate}
    \item In Brazil, mayors commonly negotiate to establish a government majority in the city council. How often do you think mayors use the following goods in exchange for support?
    \item Which of those activities do city councilors perform most often in their duties?
    \item Which of the following activities do you believe help city councilors the most in elections?
\end{enumerate}

>>>>

In this section, we show the proofs for the formal model we present in the article.

@@

\begin{proof}
To prove this result, we need to show that the mayor's expected utility satisfies the increasing differences in $g$ and $N$. This means that, when increasing the size of the council, the optimal solution $g^*$ should also increase. However, as $N$ increases discretely, we cannot take the cross-partial derivative on $g$ and $N$ or use the implicit function theorem. Instead, we use monotone comparative statics to derive these results \citep{milgrom1994monotone}.

@@

Therefore, governing costs always increase when legislature size increases.

@@

\begin{itemize}
\item \textbf{Number of Seats 2000}: Number of city councilors in 2000 (source: TSE).
\item \textbf{Population 2000}: City population according to the 2000 Brazilian Census (source: IBGE).
\item \textbf{Per Capita GDP 2000 (Millions)}: Municipal per capita GDP measured by the Brazilian Census (source: IPEA).
\item \textbf{Proportion of Poverty 2000}: Percentage of citizens earning less than R\$ 70.00 per month (source: Social Development Ministry [MDS]).
\item \textbf{Number of Seats 2004}: Number of city councilors in 2004 according to the Electoral Court (source: TSE).
\item \textbf{Infant Mortality 2005--2008}: Number of children who died before reaching one year of age, divided by the number of children born alive, multiplied by 1,000 (source: DataSUS)
\item \textbf{Postneonatal Mortality 2005--2008}: Number of children alive for at least 28 days who died before reaching one year of age, divided by the total number of children who lived for 28 days, multiplied by 1,000 (source: DataSUS)
\item \textbf{Enrollment Elementary School 2005--2008}: Number of children enrolled in elementary schools, averaged by classroom size (K--4) (source: INEP).
\item \textbf{Quality of Elementary School 2005--2008}: Average municipal IDEB scores. IDEB scores are estimated as follows. The Ministry of Education takes the average of the students' grades in math and Portuguese, then multiply the those numbers by the student yearly approval rate. INEP reweights the estimators to avoid school and classroom specific distortions (source: INEP, years: 2005, 2007).
\item \textbf{Size of Mayoral Pre-Electoral Coalition 2004}: Number of elected councilors in the mayoral pre-electoral coalition (source: TSE)
\item \textbf{Number of Appointed Bureaucrats 2005--2008}: Number of bureaucrats of the direct administration who were appointed to jobs in the municipality (source: IBGE, years: 2005, 2006, and 2008)
\item \textbf{Number of Females Elected 2004}: Number of females elected to the city council (source: TSE)
\item \textbf{Number of Non-Whites Elected 2004}: Number of non-whites elected to the city council. We collect these data for municipalities less than 10,000 inhabitants away from the cutoffs (source: own compilation based on the TSE candidate pictures).
\item \textbf{Competition per Seat 2004}: Number of people running for city councilor divided by the size of the city council (source: TSE)
\item \textbf{Proportion Approved Legislation 2005}: Number of approved legislation in 2005 divided by the number of proposed legislation (source: InterLegis)
\end{itemize}

@@

The inclusion of controls in our model has another purpose, which is to improve the efficiency of our estimates \citep{calonico2019regression}. We add five controls to the main specifications. We include the 2003 population because it conditions the multiple threshold assignments. We add GDP per capita to control for Brazil's economic expansion during the 2000s. We also add the number of seats in 2000 to ensure that our results come from the current legislature size and not from past seat allocations. Year was included in the models to control for seasonal effects. Finally, we add a dummy for Northeast states. The Northeast region received considerable attention from the Federal Government in the 2000s. According to \citet{de2015politica}, in 2003 the Lula administration recreated the Superintendency for the Development of the Northeast (SUDENE), investing BRL 84 Billion from 2003 to 2010. This amounts to around 70% of all targeted regional investments. One example of these investments is that the federal government created about 100 new universities in the Northeast region during that period.  Moreover, infant mortality in the Brazilian Northeast decreased from 22.4 deaths per 1000 infants born in 2005 to 19.6 deaths in 2008 largely due to substantial federal investments in the region. In turn, education quality improved from 2.69 in 2005 to a 3.14 score in 2007 (0 to 10 quality of education score). For this reason, adding a dummy for Northeast states is essential for ensuring the consistency and the efficiency of our estimates.

@@

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{0.9\linewidth}
      \centerline{\includegraphics[height=5in,angle=0]{../figures/simdistrRDD.pdf}}
      \caption{Simulation: RD with Multiple Thresholds -- Data Distribution}
      \bigskip
      \label{simRDdd}
    \end{minipage}
  \end{center}
\end{figure}

@@

\vspace{.2cm}

@@

For instance, changes in the controls variables substantially affect the quality of elementary education, where some sets of controls lead to a positive effect. However, most combinations of controls were insignificant.

@@

Here we evaluate the robustness of our results when we vary the number of cutoffs. We run a series of regressions limiting the population sizes to half the cutoffs, starting from the second cutoff to the last. This strategy is similar to adding one cutoff at a time. The results are stable and do not change our main conclusions. Moreover, there are no clear increasing or decreasing patterns, which would indicate the existence of differential returns. Figures \ref{senscut1} and \ref{senscut2} display the results for the main cutoffs.

@@

Two alternative explanations may confound our results. First, the mayor's selection could be affected by the size of the city council. Thus, public welfare can improve due to the quality of the mayor and not because of the partisan mechanism we describe. Second, the partisan mechanism could be capturing a dynamic similar to Weingast et al.'s [-@Weingastetal1981] "law of 1/n", whereby welfare improvements are caused by higher available revenues. This section shows that none of these concerns are valid in our case.

@@

We hand-coded 1\% of the dataset (3,466 cases) and applied a Supporting Vector Machine (SVM) algorithm to the remaining 99\% cases to classify the legislation. First, we trained and tested the accuracy of the SVM classifier on 80\% of the dataset. Then, we ran the training algorithm on the complete hand-coded data and predicted the remaining cases.\footnote{We tested SVM, Naïve Bayes, Random Forests, and Neural Networks for this task. We have chosen SVM as it gives the highest prediction rate of all of the algorithms we have tried. We use a simple bag-of-words classifier provided by the \texttt{RTextTools} package for \texttt{R} \citep{collingwood2013rtexttools}.}

@@

To weight the survey, we use the Legislative Census, conducted by *Interlegis* on the behalf of the Brazilian Senate. The company was hired to improve the quality of Brazilian government branches. We include the following categories in the weighting procedure: region of Brazil, legislature size (9 to 15 or more), gender, and age less than 39. We use the proportions in each bin to weight the graphs and statistics generated after the raking process.\footnote{The \texttt{survey} package for \texttt{R} describes the raking process in more detail \citep{lumley2004analysis}. In summary, it iterates the post-stratification procedure until the sample marginals match the population marginals for all variables.} Table \ref{tabweight} shows the sample, population, and weighted proportions.

@@

\begin{enumerate}
    \item In Brazil, mayors usually negotiate to establish a government majority in the city council. How often do you think mayors use the following goods in exchange for support?
    \item Which of those activities do city councilors perform most often in their duties?
    \item Which of the following activities do you believe helps city councilors the most in elections?
\end{enumerate}
